Syntax Is All You Need: A Universal-Language Approach to Mutant Generation

### 2023-03-29 paper/main.tex
Adding Rust numbers
broken labels, mihnor spacing on table 3

Tables~\ref{tab:table_cpp1}-\ref{tab:table_rust1} compare generated
mutants for Regex and Comby modes of our tool.  Invalid
mutants are mutants that the tool produces, but which fail to
compile.  The cost of each such mutant is usually small,
since compilation generally fails early for trivial reasons (e.g.,
{\tt break} or {\tt continue} outside a loop), 
but reducing the number of such mutants directly reduces the cost of
initial mutant generation, and generally indicates a better
``understanding'' of the source being mutated.

% ---------------- C++ table UM Only--------------------------

@@

\label{tab:table_cpp1java}

@@

libgraphqlparser& 137 & 741 & 21.19\%  & 440 & 29.77\% \\
rethinkdb\_rebirth &471 & 2246 & 24.27\% & 2009 & 24.49\% \\

@@

\end{tabularx}}

@@

\begin{table}[hbtp]
\centering
{\footnotesize % feels bad man but it's definitely easier than most other alternatives
\caption{Regex vs. Comby for Python.   LOC describes the project files considered. Muts. is total number of mutants generated for those files. \claire{FIXME: what's an invalid mutant for python?}  We can use TCE to check for redundant Python mutants, shown under ``Red.''}
\label{tab:table_python1}
\begin{tabularx}{\columnwidth}{l|r|rcr|rcr}
\toprule\toprule
                 &                 & \multicolumn{3}{c|}{\textbf{Regex}}             &\multicolumn{3}{c}{\textbf{Comby}}  \\[1ex]
\textbf{Project} & \textbf{LOC}    & \textbf{Muts.} & \textbf{Red.}  & \textbf{\% Valid} & \textbf{Muts.} & \textbf{Red.}   &  \textbf{\% Valid} \\\midrule
keract           & 402             & 3984  & 729   & 41.53\%           & 3067  & 533    & 47.96\% \\
dtw              & 118             & 2106  & 567   & 46.20\%           & 1835  & 639    & 47.19\% \\
notion-          & \mr{2}{189}     & \mr{2}{1038}  & \mr{2}{82}    & \mr{2}{18.89\%}           & \mr{2}{1019}  & \mr{2}{122}    & \mr{2}{19.23\%} \\
sdk-py           & & & & & & \\[0.5ex]
atlassian-       & \mr{2}{236}      & \mr{2}{1541}  & \mr{2}{207}   & \mr{2}{19.99\%}           & 1319  & \mr{2}{351}    & \mr{2}{25.25\%} \\
python-api       & & & & & & \\[0.5ex]
ESD              & 698             & 6268  & 1642  & 29.46\%           & 3874  & 614    & 44.18\% \\
fiber            & 358             & 1565  & 163   & 31.08\%           & 1466  & 241    & 31.24\% \\\bottomrule
\end{tabularx}
\begin{tabularx}{\columnwidth}{Xllrrrrrr}
                  && &  \textbf{Red.}     & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
\multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
    &   &  Std.Dev. & 584.79  & 708.93 & 813.35 & 11.06\%  \\
 & \multirow{2}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
  &   &  Std.Dev. &  211.62 & 628.79 & 418.99 & 12.28\%   \\\bottomrule
\end{tabularx}}
\end{table}

@@

\begin{table}[htbp]
\centering
\caption{Java (Our Implementation vs. PIT vs. LittleDarwin). \claire{FIXME: explain why no eqv for PIT.} LittleDarwin produces no equivalent mutants.\claire{right?}}
\label{tab:table_java2}
\resizebox{\columnwidth}{!}{% again don't love it, but footnotesize is actually smaller than necessary
\begin{tabular}{l|rrr|rr|rr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{2}{c|}{\textbf{PIT}} & \mc{2}{c}{\textbf{LittleDarwin}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} \\ \midrule

@@

\begin{table*}[htbp]
\centering
{
\caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
\label{tab:table_python2}
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{3}{c|}{\textbf{Mutmut}} & \mc{3}{c}{\textbf{CosmicRay}} \\
\textbf{Project} & \textbf{Muts} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} 
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
\midrule
keract & 1471 & 0.26 & 10\% & 482 & 0.72 & 10\% & 921 & 0.27 & 10\% \\
dtw & 866 & 0.55 & 10\% & 270 & 0.44 & 0\% & 713 & 0.46 & 20\% \\
notion-sdk-py & 196 & 0.60 & 20\% & 63 & 0.65 & 0\% & 52 & 0.44 & 0\% \\
atlassian-python-api & 333 & 0.62 & 40\% & 174 & 0.71 & 10\% & 106 & 0.74 & 10\% \\
ESD & 1712 & 0.07 & 10\% & 720 & 0.08 & 0\% & 1013 & 0.06 & 10\% \\
Fiber & 458 & 0.64 & 20\% & 222 & 0.57 & 40\% & 320 & 0.81 & 0\% \\\midrule
\textbf{Mean} &  839.33 & 0.46 & 18.00\% & 321.83 & 0.53 & 10.00\% & 520.93 & 0.46 & 8.33\% \\
\textbf{Std.Dev.} &  628.80 & 0.24 & 11.69\% & 239.04 & 0.24 & 15.49\% & 417.50 & 0.28 & 7.53\% \\\bottomrule        
\end{tabular}
}
\end{table*}

% ---------------- Rust table (UM vs Others)--------------------------

\begin{table}[htbp]
\centering
\caption{Rust (Our Implementation vs. Cargo-Mutants)}
\label{tab:table_rust2}
\resizebox{\columnwidth}{!}{
\begin{tabular}{|c|c|r|r|r|r|}
\hline
 &  & \textbf{Killed} & \textbf{Not Killed} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{Equivalent \%} \\ \hline

 \multirow{3}{*}{\small cargo release} & Regex & 721 & 130 & 0.85 & 0\% \\\cline{2-6}
    & Comby & 1344 & 346 & 0.80 & 0\% \\\cline{2-6}
    & Cargo-Mutants & 2	& 3 & 0.40 & 0\% \\ \hline
\multirow{3}{*}{\small passerine} & Regex & 122 & 91 & 0.57 & 20\% \\\cline{2-6}
    & Comby & 90 & 83 & 0.52 & 50\% \\ \cline{2-6}
    & Cargo-Mutants & 0	& 0 & 0 & 0\% \\ \hline
\multirow{3}{*}{\small typos} & Regex & 295 & 91 & 0.76 & 40\% \\\cline{2-6}
    & Comby & 283 & 83 & 0.77 & 20\%\\ \cline{2-6}
    & Cargo-Mutants & 5 & 1 & 0.83 & 0\%\\ \hline

\multirow{3}{*}{\small ord} & Regex & 558 & 204 & 0.73 & 10\% \\\cline{2-6}
    & Comby & 533 & 110 & 0.83 & 20\% \\ \cline{2-6}
    & Cargo-Mutants & 6 & 0 & 1.00 & 0\% \\ \hline
\multirow{3}{*}{\small bazuka} & Regex & 281 & 170 & 0.62 & 40\% \\\cline{2-6}
    & Comby & 291 & 111 & 0.73 & 0\% \\ \cline{2-6}
    & Cargo-Mutants & 0 & 0 & 0 & 0\% \\ \hline
\multirow{3}{*}{\small strum} & Regex & 8 & 43 & 0.16 & 10\%\\\cline{2-6}
    & Comby & 11 & 39 & 0.22 & 40\%\\ \cline{2-6}
    & Cargo-Mutants & 2 & 0 & 1.00 & 0\%\\ \hline

\multirow{9}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.62  & 20.00\%\\\cline{3-6}
    &   &  \multicolumn{2}{c|}{Median} & 0.68 & 15.00\%\\\cline{3-6}
    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & 16.73\% \\\cline{2-6}

 & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.64 & 22.00\%\\\cline{3-6}
    &   &  \multicolumn{2}{c|}{Median} & 0.75 & 20.00\% \\\cline{3-6}
    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.25 & 20.41\% \\\cline{2-6}

 & \multirow{3}{*}{Cargo-Mutants} & \multicolumn{2}{c|}{Mean} & 0.54 & 0.00\%\\\cline{3-6}
    &   &  \multicolumn{2}{c|}{Median} & 0.62 & 0.00\% \\\cline{3-6}
    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.47 & 0.00\%\\ \hline

\begin{table}[htbp]
\centering
{\small
\caption{Rust (Our Implementation vs. Cargo-Mutants)}
\label{tab:table_rust2}
\begin{tabularx}{\columnwidth}{X|crr|crr}
\hline
        & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Cargo-Mutants} \\ 
Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
\hline
cargo release & \\

\textbf{Mean}   &  \\
\textbf{Std. Dev} &  \\
\end{tabularx}
}
\end{table}

@@

Our tools produce somewhat lower mutation scores than other tools, and
higher (though still generally good) equivalence rates than some
tools, but clearly fall into the same broad similarity of behavior
reported for high-quality mutation tools in the literature comparing
Java mutation tools.   The tools with very low or zero equivalence rates, furthermore,
are very conservative in mutation generation, and often simply don't
produce many mutants.  In fact, inspection of the non-comment source
lines as measured by CLOC shown in Tables~\ref{tab:table_java1} and
\ref{tab:table_java2} and Tables~\ref{tab:table_rust1} and
\ref{tab:table_rust2} shows that for many programs, LittleDarwin and
Cargo-Mutants generated \emph{much less than one mutant per actual
  line code}.  Such a conservative approach is likely to produce few
or no actionable (unkilled and non-equivalent) mutants for a program
with a very good, high-coverage test suite.  But such suites are
arguably the ones where mutation testing offers the most promise, by
allowing the detection of very subtle remaining holes in tests for
high-criticality software.

>>>>

Table~\ref{tab:table_cpp1javarust} and Table~\ref{tab:table_python1} compare generated
mutants for Regex and Comby modes of our tool.  Invalid
mutants are mutants that the tool produces, but which fail to
compile.  The cost of each such mutant is usually small,
since compilation generally fails early for trivial reasons (e.g.,
{\tt break} or {\tt continue} outside a loop), 
but reducing the number of such mutants directly reduces the cost of
initial mutant generation, and generally indicates a better
``understanding'' of the source being mutated.

% ---------------- C++/Java/Rust table UM Only--------------------------

@@

\label{tab:table_cpp1javarust}

@@

{\small libgraphqlparser} & 137 & 741 & 21.19\%  & 440 & 29.77\% \\
{\small rethinkdb\_rebirth} &471 & 2246 & 24.27\% & 2009 & 24.49\% \\

@@

@@

\begin{table}[hbtp]
\centering
{\footnotesize % feels bad man but it's definitely easier than most other alternatives
\caption{Regex vs. Comby for Python.   LOC describes the project files considered. Muts. is total number of mutants generated for those files. \clg{FIXME: what's an invalid mutant for python?}  We can use TCE to check for redundant Python mutants, shown under ``Red.''}
\label{tab:table_python1}
\begin{tabularx}{\columnwidth}{l|r|rcr|rcr}
\toprule\toprule
                 &                 & \multicolumn{3}{c|}{\textbf{Regex}}             &\multicolumn{3}{c}{\textbf{Comby}}  \\[1ex]
\textbf{Project} & \textbf{LOC}    & \textbf{Muts.} & \textbf{Red.}  & \textbf{\% Valid} & \textbf{Muts.} & \textbf{Red.}   &  \textbf{\% Valid} \\\midrule
keract           & 402             & 3984  & 729   & 41.53\%           & 3067  & 533    & 47.96\% \\
dtw              & 118             & 2106  & 567   & 46.20\%           & 1835  & 639    & 47.19\% \\
notion-          & \mr{2}{189}     & \mr{2}{1038}  & \mr{2}{82}    & \mr{2}{18.89\%}           & \mr{2}{1019}  & \mr{2}{122}    & \mr{2}{19.23\%} \\
sdk-py           & & & & & & \\[0.5ex]
atlassian-       & \mr{2}{236}      & \mr{2}{1541}  & \mr{2}{207}   & \mr{2}{19.99\%}           & 1319  & \mr{2}{351}    & \mr{2}{25.25\%} \\
python-api       & & & & & & \\[0.5ex]
ESD              & 698             & 6268  & 1642  & 29.46\%           & 3874  & 614    & 44.18\% \\
fiber            & 358             & 1565  & 163   & 31.08\%           & 1466  & 241    & 31.24\% \\\bottomrule
\end{tabularx}
\begin{tabularx}{\columnwidth}{Xllrrrrrr}
                  && &  \textbf{Red.}     & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
\multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
    &   &  Std.Dev. & 584.79  & 708.93 & 813.35 & 11.06\%  \\
 & \multirow{2}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
  &   &  Std.Dev. &  211.62 & 628.79 & 418.99 & 12.28\%   \\\bottomrule
\end{tabularx}}
\end{table}

@@

\begin{table}[htbp]
\centering
\caption{Java (Our Implementation vs. PIT vs. LittleDarwin). \clg{FIXME: explain why no eqv for PIT.} LittleDarwin produces no equivalent mutants.\clg{right?}}
\label{tab:table_java2}
\resizebox{\columnwidth}{!}{% again don't love it, but footnotesize is actually smaller than necessary
\begin{tabular}{l|rrr|rr|rr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{2}{c|}{\textbf{PIT}} & \mc{2}{c}{\textbf{LittleDarwin}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} \\ \midrule

@@

\begin{table*}[htbp]
\centering
{
\caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
\label{tab:table_python2}
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{3}{c|}{\textbf{Mutmut}} & \mc{3}{c}{\textbf{CosmicRay}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} 
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
\midrule
keract & 1471 & 0.26 & 10\% & 482 & 0.72 & 10\% & 921 & 0.27 & 10\% \\
dtw & 866 & 0.55 & 10\% & 270 & 0.44 & 0\% & 713 & 0.46 & 20\% \\
notion-sdk-py & 196 & 0.60 & 20\% & 63 & 0.65 & 0\% & 52 & 0.44 & 0\% \\
atlassian-python-api & 333 & 0.62 & 40\% & 174 & 0.71 & 10\% & 106 & 0.74 & 10\% \\
ESD & 1712 & 0.07 & 10\% & 720 & 0.08 & 0\% & 1013 & 0.06 & 10\% \\
Fiber & 458 & 0.64 & 20\% & 222 & 0.57 & 40\% & 320 & 0.81 & 0\% \\\midrule
\textbf{Mean} &  839.33 & 0.46 & 18.00\% & 321.83 & 0.53 & 10.00\% & 520.93 & 0.46 & 8.33\% \\
\textbf{Std.Dev.} &  628.80 & 0.24 & 11.69\% & 239.04 & 0.24 & 15.49\% & 417.50 & 0.28 & 7.53\% \\\bottomrule        
\end{tabular}
}
\end{table*}

\begin{table}[htbp]
\centering
{\small
\caption{Rust (Our Implementation vs. Cargo-Mutants)}
\label{tab:table_rust2}
\begin{tabularx}{\columnwidth}{X|rrr|rrr}
\toprule
        & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{3}{c}{\textbf{Cargo-Mutants}} \\ 
\textbf{Project} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\\midrule
cargo release & 1690 & 0.80 & 0\% & 5 & 0.40 & 0\%   \\
passarine & 173 & 0.52 & 50\% & 0 & 0.00 & 0\%   \\
typos & 366 & 0.77 & 20\% & 6 & 0.83 & 0\%   \\
ord & 643 & 0.83 & 20\% & 6 & 1.00 & 0\%   \\
bazuka & 402 & 0.73 & 0\% & 0 & 0.00 & 0\%   \\
strum & 50 & 0.22 & 40\% & 2 & 1.00 & 0\%   \\\midrule
\textbf{Mean}   & 554.00 & 0.64 & 22.00\% & 3.17 & 0.54 & 0.00\%\\
\textbf{Std. Dev} & 592.63 & 0.25 & 20.41\%   & 2.86 & 0.47 & 0.00\% \\\bottomrule
\end{tabularx}
}
\end{table}

@@

Our tools produce somewhat lower mutation scores than other tools, and
higher (though still generally good) equivalence rates than some
tools, but clearly fall into the same broad similarity of behavior
reported for high-quality mutation tools in the literature comparing
Java mutation tools.   The tools with very low or zero equivalence rates, furthermore,
are very conservative in mutation generation, and often simply don't
produce many mutants.  In fact, inspection of the non-comment source
lines as measured by CLOC as compared to results as shown in Tables~\ref{tab:table_cpp1javarust},
Tables~\ref{tab:table_java2} and 
\ref{tab:table_rust2} shows that for many programs, LittleDarwin and
Cargo-Mutants generated \emph{much less than one mutant per actual
  line code}.  Such a conservative approach is likely to produce few
or no actionable (unkilled and non-equivalent) mutants for a program
with a very good, high-coverage test suite.  But such suites are
arguably the ones where mutation testing offers the most promise, by
allowing the detection of very subtle remaining holes in tests for
high-criticality software.
